`r opts_chunk$set(cache=TRUE)`

The following example aimed to explore variability in posterior densities based on different random samples and for three different sample sizes.

$latex y_1 \sim N(\mu_{y1}, \sigma^2_{y1})$
$latex y_2 \sim N(\mu_{y2} + \beta_{y2}, \sigma^2_{y2})$



# Import packages
```{r message=FALSE}
rm(list=ls())
library(rjags)
library(coda)
library(psych)
library(plyr)
```


# Generate Data
```{r}
simulatedata <- function(N=200, MuY1=1, BetaY2=2, SigmaY1=1) {
    Y1 <- rnorm(N, MuY1, SigmaY1)
    Y2 <- rnorm(N, MuY1 + BetaY2, SigmaY1)
    
    list(Y1=Y1, Y2=Y2)
}    
```

# Simple analyses
```{r}
Data <- simulatedata()
plot(Data$Y1, Data$Y2, ylim=c(1,5), xlim=c(1,5)); abline(a=0, b=1)
```

# JAGS Model
```{r JAGSmodel, tidy=FALSE}

modeltext <- "
model {
    # Model
    for (i in 1:length(Y1)) {
        Y1[i] ~ dnorm(MuY1, TauY1)
        Y2[i] ~ dnorm(MuY1 + BetaY2, TauY1)
    }

    # Priors
    MuY1  ~ dnorm(0, 1.0E-3)
    BetaY2  ~ dunif(0, 1000)
    TauY1 ~ dgamma(1.0E-3, 1.0E-3);
    SDY1 <- TauY1 ^ -0.5
}
"
```
    


# Fit Jags Models
```{r}
fitmodel <- function() {
    mod1 <- jags.model(textConnection(modeltext), data=Data, n.chains=1, n.adapt=500, quiet=TRUE)
    update(mod1, 500) # burn in
    mod1.samples <- coda.samples(model=mod1, n.iter=1000, thin=1,
                                 variable.names=c('MuY1', 'BetaY2', 'SDY1'))
    summary(mod1.samples)$statistics[,1]
}
```

# Setup iterative simulator
```{r message=FALSE}
Iterations <- 50 # put it up to 200 later

runs <- expand.grid(N=c(5, 20, 100), iteration=seq(Iterations))
runs$BetaY2 <- NA
runs$MuY1 <- NA
runs$SDY1 <- NA

for (i in seq(nrow(runs))) {
    Data <- simulatedata(N=runs[i, "N"])    
    fits <- fitmodel()
    runs[i, names(fits)] <- fits
}
```

The simulation is run `r Iterations` times for each selected samples size (5, 20, and 100).

For each run:

1. Data is simulated based on the specification (e.g., sample size).
2. The posterior mean of each parameter is extracted.
3. The values for the run are stored in a dataframe, one row per run.


# Examine simulation results
```{r}
par(mfrow=c(3,1))
boxplot(MuY1 ~ N, runs, main = "MuY1"); abline(h=1)
boxplot(BetaY2 ~ N, runs, main ="BetaY2" ); abline(h=2)
boxplot(SDY1 ~ N, runs, main="SDY1"); abline(h=1)


results <- ddply(runs, .(N), summarise, 
      mean.MuY1=mean(MuY1),
      mean.BetaY2=mean(BetaY2),
      mean.SDY1=mean(SDY1),
      sd.MuY1=sd(MuY1),
      sd.BetaY2=sd(BetaY2),
      sd.SDY1=sd(SDY1),
      min.MuY1=min(MuY1),
      min.BetaY2=min(BetaY2),
      min.SDY1=min(SDY1),
      max.MuY1=max(MuY1),
      max.BetaY2=max(BetaY2),
      max.SDY1=max(SDY1))

t(round(results, 4))
```
