r opts_chunk$set(cache=TRUE)

Data are observed on observations $y_i$, where $latex i=1, \ldots, I$.
Data are normally distributed with mean $latex \mu_i$ and constant variance.

$$latex y_i \sim N(\mu_i, \sigma^2)$$

$latex \mu_i$ is equal to the mean of the $latex \gamma_j$, the mean of the $latex j$th latent group.

$$latex \mu_i = \sum_{j=1}^3 \beta_{ji} \gamma_j$$

$latex \beta_{ji}$ indicates whether individual $latex i$ is a member of the $latex j$th latent group.

$$latex \beta_{ji} = \begin{cases} 
    1 & \text{if $x=j$,}    \\
    0 &\text{otherwise}
\end{cases}$$

$$latex x_i \sim \text{Cat}(\pi)$$

The priors on the probability of membership of each of the three latent groups are

$$latex \pi_1 \sim \text{Unif}(0, 1)$$
$$latex \pi_2 \sim \text{Unif}(0, 1 - \pi_1)$$
$$latex \pi_3 = 1 - \pi_1 - \pi_2.$$

$latex \pi_3$ is constrained so that once $latex \pi_1$ and $latex \pi_2$ are known, the probabilities must sum to 1.0. 

These are the priors on the means of each of the latent groups.

$$latex \gamma_1 \sim \text{Unif}(-100, 100)$$
$$latex \gamma_2 \sim \text{Unif}(\gamma_1, 100)$$
$$latex \gamma_3 \sim \text{Unif}(\gamma_2, 100)$$
The priors are structured to ensure that an order is maintained such that $latex \gamma_1$ is the lowest and $latex \gamma_3$ is the highest.

An uninformative prior is placed on precision:

$$latex \tau = \frac{1}{\sigma^2}$$
$$latex \tau \sim \text{Unif}(0, 1000)$$



# Setup
```{r setup, message=FALSE}
rm(list=ls())
require(rjags)
require(coda)
```


# Generate data
```{r generate_data}
N <- 100

x <- sample(1:3, replace=TRUE, size=N, prob=c(.2, .3, .5))
Beta1 <- as.numeric(x == 1)
Beta2 <- as.numeric(x == 2)
Beta3 <- as.numeric(x == 3)
Gamma <- c(0,5,10) # latent group means

Mu <- Beta1 * Gamma[1] + 
      Beta2 * Gamma[2] + 
      Beta3 * Gamma[3]
Sigma <- 1
y <- rnorm(N, Mu, Sigma)

Data <- list(y=y)
```

# Basic data analysis
```{r basic_analyses}
par(mfrow=c(1,2))
plot(density(y))
boxplot(y ~ x)
```

# Specify JAGS model
```{r jagsmodel, tidy=FALSE}
jags.script <- "
model{
    # likelihood
    for( i in 1:length(y)) {
        x[i] ~ dcat(Pi)
        Beta1[i] <- x[i] == 1
        Beta2[i] <- x[i] == 2
        Beta3[i] <- x[i] == 3
        Mu[i] <- Beta1[i] * Gamma[1] + 
            Beta2[i] * Gamma[2] + 
            Beta3[i] * Gamma[3]
        y[i] ~ dnorm(Mu[i], Tau)
    }

    # priors
    Pi[1] ~ dunif(0, 1)
    Pi[2] ~ dunif(0, 1-Pi[1])
    Pi[3] <- 1 - Pi[1] - Pi[2]

    Gamma[1] ~ dunif(-100, 100)
    Gamma[2] ~ dunif(Gamma[1], 100)
    Gamma[3] ~ dunif(Gamma[2], 100)
    Tau ~ dunif(0, 1000)    
    Sigma <- Tau ^ -0.5
}
"
```

# Perform Bayesian analysis using JAGS
```{r JAGSanalysis}
jags.fit <- jags.model(textConnection(jags.script), data=Data,
                       n.chains=4, n.adapt=1000)

update(jags.fit, n.iter=1000) # burnin

jags.samples <- coda.samples(model=jags.fit, variable.names=c("Gamma", "Sigma"), n.iter=1000)
                             
plot(jags.samples) # plot trace and posterior density for each parameter
(summary.jags.model <- summary(jags.samples))
```

```{r compare_jags_with_true}
cbind(True=c(Gamma, Sigma), 
    round(summary.jags.model$statistics[,1:2], 3),
    round(summary.jags.model$quantiles[,c(1,5)], 3))
```
