`r opts_chunk$set(cache=TRUE)`
<!-- Fit normal distribution to some data -->

The following post shows an example of fitting a normal distribution with unknown mean and variance to data.

<!-- more -->
Observed data are $latex y_i$ where $latex i=1, \ldots, I$.  The data is modelled as independent observations drawn from a normal distribution with unknown variance:
$$latex y_i \sim N(\mu, \sigma^2)$$

The inverse of the variance $latex \sigma^2$ is the precision $latex \tau$. 

$$latex \tau = \frac{1}{\sigma^2}$$

Noninformative priors are placed on $latex \mu$ and $latex \tau$

$$latex \mu \sim N(0, 0.001)$$
$$latex \tau \sim \Gamma(0.001, 0.001)$$


# Import packages
```{r message=FALSE}
rm(list=ls())
library(rjags)
library(coda)
library(ggplot2)
```


## Simulate data
```{r}
set.seed(4444)
N <- 100 # number of subjects
Mu <- 100
Sigma <- 15 # standard deviation of error

y <- rnorm(n=N, mean=Mu, sd=Sigma)
Data <- data.frame(y)

# Convert data to input format required by JAGS
jagsdata <- list(y=y)
```


## Traditional analysis
```{r fig.width=3, height=3}
ggplot(Data, aes(x=y)) + geom_density(fill="lightblue", alpha=.5)
c(n=length(y), 
  mean=mean(y), 
  sd=sd(y),
  se.mean=sd(y)/sqrt(length(y)))
```

# Specify and export BUGS model
```{r}
jags.script <- "
model {
    # Model
    for (i in 1:length(y)) {
        y[i]  ~ dnorm(mu, tau)
    }
    
    # Priors
    mu  ~ dnorm(0, 0.001);
    tau ~ dgamma(0.001, 0.001);
    
    # Transformations
    sigma  <- 1.0/sqrt(tau);
}
"
```


## Bayesian analysis
```{r}
mod1 <- jags.model(textConnection(jags.script), data=jagsdata, n.chains=4, n.adapt=1000)
update(mod1, 200) # burn in

# monitor
mod1.samples <- coda.samples(model=mod1,
                             variable.names=c('mu', 'sigma'),
                             n.iter=1000)                  
```


## Model summary and diagnostics
```{r}
plot(mod1.samples) # plot trace and posterior density for each parameter
summary(mod1.samples) # print descriptive statistics of posterior densities for parameters
```

